Inferring the Recombinant Genealogy of a Pandemic Pathogen in Real Time

### 2025-09-15 latex/main.tex

\subsection{Analysis of Pango lineage ``origination'' events}
\label{sec:supplement_pango_lineages}

To analyze how well the ARG reflects the phylogenetic structure implicit in the
Pango lineage naming system, we considered the Pango assignments generated by
Pangolin on the alignments for each node in the ARG (STAR Methods). If the ARG
perfectly reflected the Pango lineage structure, each lineage would form a
clade descending from a single originating node with that label. We identified
putative origination nodes among those labeled with a given lineage as the node
with the maximum number of descendant samples and the earliest time (if there
is a tie).

Of the 2058 distinct Pango lineages in the ARG, 1469 of these (comprising
743922 samples) match perfectly, with unique origination events in the ARG
where all samples assigned a given lineage descend from the first node assigned
to that lineage. A further 310 lineages (398391 samples) match perfectly when
we count the descendants of the parent of the first node (accounting for
polytomies in which multiple originating nodes for a given lineage are
siblings). We then have 255 lineages (996490 samples) where the difference in
the number of descendants of the first node's parent is $<$ 100. The remaining 25
lineages (344079) are dominated by a few large lineages such as BA.1.1 (155595
samples) and AY.4.2 (54607 samples) which have multiple non-sibling origins
within the ARG.

\subsection{Additional analyses of the Pango Xs}

>>>>

\subsection{Analysis of Pango lineage ``origination'' events}
\label{sec:supplement_pango_lineages}
To analyze how well the ARG reflects the phylogenetic structure implicit in the
Pango lineage naming system, we considered the Pango assignments generated by
Pangolin on the alignments for each node in the ARG (STAR Methods). If the ARG
perfectly reflected the Pango lineage structure, each lineage would form a
clade descending from a single originating node with that label. We identified
putative origination nodes among those labeled with a given lineage as the node
with the maximum number of descendant samples and the earliest time (if there
is a tie).

Of the 2058 distinct pango lineages in the ARG, 1473 of these (comprising
717798 samples) match perfectly, with unique origination events in the ARG
where all samples assigned a given lineage descend from the first node assigned
that lineage. A further 245 lineages (589197 samples) match perfectly when we
count the descendants of the parent of the first node (accounting for
polytomies in which multiple originating nodes for a given lineage are
siblings). We then have 306 lineages (755840 samples) where the difference in
the number descendants of the first node's parent is $<$ 100. The remaining 35
lineages (420047 samples) are dominated by a few large lineages such as BA.1.1
(147271 samples) and AY.4.2 (54607 samples) which have multiple non-sibling
origins within the ARG.

See notebook
\href{https://github.com/jeromekelleher/sc2ts-paper/blob/main/notebooks/analysis\_pango\_events\_in\_arg.ipynb}
{analysis\_pango\_events\_in\_arg.ipynb}
for details.

\subsection{Additional analyses of the Pango Xs}
\label{sec:supplement_indepth_pango_x_lineages}
